# The Most Harmful Weather Event in US
---
## 1. Synopsis
In US, storm and other weather events cause a large loss for both population 
health and economy every year.In order to reduce the loss and damage from these
disaster, it is important to find out which of them are the most harmful.
History information including time and geography data for each weather event occurrence data were collected by U.S. National Oceanic and Atmospheric 
Administration's (NOAA) storm database. In this paper, the database was 
downloaded and analysed to answer two question:

1. Across the United States, which types of events (as indicated in the EVTYPE 
variable) are most harmful with respect to population health?
2. Across the United States, which types of events have the greatest economic 
consequences?

## 2. Getting the data
Dataset is downloaded from coursera web page

```{r download.data, cache=TRUE}
Sys.setlocale('LC_ALL', 'English')
download.file("http://d396qusza40orc.cloudfront.net/repdata/data/StormData.csv.bz2", 
              "repdata-data-StormData.csv.bz2")
stormData <- read.csv("repdata-data-StormData.csv.bz2", sep=",", header=TRUE, 
                      stringsAsFactors=FALSE)
```

Take a look at the dataset, there are 4 columns are corresponding to health and economy loss:

1. FATALTIES ~ The number of fatalities caused by the events
2. INJURIES ~ The number of people injured by the events
3. PROPDMG ~ The amount of property loss by the events
4. CROPDMG ~ The amount of crop damage by the weather

```{r}
str(stormData)
```

## 3. Data Processing
### 3.1 The most harmful events for public health
Top ten health threaten events ranked by total fataliteis and injuries.
```{r}
fatalities <- aggregate(stormData$FATALITIES, list(stormData$EVTYPE), sum)
injuries <- aggregate(stormData$INJURIES, list(stormData$EVTYPE), sum)
health <- merge(fatalities, injuries, by="Group.1")
health[,4] <- health[,2] + health[,3]
colnames(health) <- c("type", "fatalities", "injuries", "total")
healthtop <- health[order(health$total, decreasing=TRUE),][1:10,]
```

Plot the data
```{r}
library(reshape2)
healthtop_1 <- melt(healthtop, id.vars="type")

library(ggplot2)
ggplot(healthtop_1, aes(x = reorder(type, -value), y = value)) + 
    geom_bar(stat = "identity", aes(fill = variable), position = "dodge") + 
    scale_y_sqrt("Number of Cases") + xlab("Event Type") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) + 
    ggtitle("Top 10 Health Threat Events")
```

### 3.2 The most harmful events for economy
By examing the data, we find that there are huge difference between the range of  property damage and corp damage. So it is more 
```{r}
propertydmgs <- aggregate(stormData$PROPDMG, list(stormData$EVTYPE), sum)
corpdmgs <- aggregate(stormData$CROPDMG, list(stormData$EVTYPE), sum)
colnames(propertydmgs) <- c("type", "propertydmgs")
colnames(corpdmgs) <- c("type", "corpdmgs")

propertydmgstop <- propertydmgs[order(propertydmgs$propertydmgs, 
                                      decreasing=TRUE),][1:10,]
corpdmgstop <- corpdmgs[order(corpdmgs$corpdmgs, decreasing=TRUE),][1:10,]
```

```{r}
library(gridExtra)

plot2 <- ggplot(propertydmgstop, aes(x=reorder(type, -propertydmgs), 
                                     y=propertydmgs)) + 
    geom_bar(stat = "identity",  position = "dodge") + 
    scale_y_sqrt("Property Damage(US Dollar)") + xlab("Event Type") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) + 
    ggtitle("Top 10 Property Threat Events")

plot3 <- ggplot(corpdmgstop, aes(x=reorder(type, -corpdmgs), y=corpdmgs)) + 
    geom_bar(stat = "identity",  position = "dodge") + 
    scale_y_sqrt("Crop Damage(US Dollar)") + xlab("Event Type") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) + 
    ggtitle("Top 10 Crop Threat Events")

grid.arrange(plot2, plot3, ncol = 2)
```

### 3.3 Which part of US suffer most from the weather disaster?

## 4. Result 
According to the 